---
title: "Plant trait cleaning and combining"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant trait data by synonyms species list and combine."
data input: "TropicalAndes_all_plant_traits_harmonized_subset.csv"
data output: "ADD"
date: "2023-11-04"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(rstatix)
library(funbiogeo)
library(visdat)
```

# Set file paths
```{r}
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Read in subset fruiting species data
```{r}
plant_traits <- read.csv(file.path(data_path_L1, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
```

# Look at trait numbers
```{r}
plant_traits %>%
  count(TraitName)
```

```{r}
plant_traits %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

Final trait list:
- dispersal syndrome
- Fruit mass
- Fruit length
- Plant growth form
- Plant height
- Plant Lifespan
- Seed mass
- Seed length
- Seed width
- Fruit type


# Standardize trait measurements
## Create functions
```{r}
numeric_data_average <- function(data, unit, traitname) {
  # Remove rows if TraitValue is non-numeric and convert to numeric
  data <- data %>%
    filter(!is.na(as.numeric(TraitValue))) %>%
    mutate(TraitValue = as.numeric(TraitValue))
  
  # Remove duplicate rows
  data <- data %>%
    distinct(.keep_all = TRUE)
  
  # Remove rows with missing or empty species names
  data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
  
  # Calculate geometic mean value, count, and variance per trait per species
  summary_data <- data %>%
    group_by(Accepted_species) %>%
    summarise(
      TraitValue_mean = exp(mean(log(TraitValue))),
      records_used = n(),
      variance = var(TraitValue, na.rm = TRUE)
    )
  
  # Add trait name column
  summary_data$TraitName <- traitname
  
  # Rename the mean column to TraitValue
  summary_data <- summary_data %>%
    rename(TraitValue = TraitValue_mean)
  
  return(summary_data)
}

```

```{r}
factor_data_merge <- function(data, traitname) {
  
  # Remove rows with missing or empty species names
  data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
  
  # Subset data to just species and TraitValue
  data <- data[, c("Accepted_species", "TraitValue")]
  
  # Calculate mode, count, and variance-like measure (using mode frequency)
  summary_data <- data %>%
    group_by(Accepted_species) %>%
    summarise(
      ModeValue = names(sort(table(TraitValue), decreasing = TRUE))[1],
      records_used = n(),
      mode_freq = max(table(TraitValue)),
      variance = 1 - (max(table(TraitValue)) / n())
    )
  
  # Add trait name column
  summary_data$TraitName <- traitname
  
  # Rename the mean column to TraitValue
  summary_data <- summary_data %>%
    rename(TraitValue = ModeValue)
  
  return(summary_data)
}
```

```{r}
# Detect outliers and report species with multiple entries per trait
detect_outliers <- function(data) {
  outlier_results <- data %>%
    group_by(Accepted_species) %>%
    filter(n() >= 3) %>%
    nest() %>%
    mutate(outliers = map(data, ~identify_outliers(.x, variable = "TraitValue"))) %>%
    unnest(outliers) %>%
    select(Accepted_species, TraitName, TraitValue, is.outlier, is.extreme) %>%
    ungroup()
  
  return(outlier_results)
}

# Get the species with outliers
species_with_outliers <- function(data) {
  outliers_by_species <- data %>%
    group_by(Accepted_species) %>%
    summarize(outlier_count = sum(is.outlier, na.rm = TRUE)) %>%
    filter(outlier_count > 0) %>%
    ungroup()
  
  return(outliers_by_species)
}
```

```{r}
summarize_species_records <- function(df) {
  # Count the total number of species
  total_species <- df %>% 
    distinct(Accepted_species) %>% 
    nrow()
  
  # Count the number of species with only one trait record
  species_with_one_record <- df %>% 
    group_by(Accepted_species) %>% 
    summarize(record_count = n()) %>% 
    filter(record_count == 1) %>% 
    nrow()
  
  # Count the number of species with more than three records
  species_with_more_than_three_records <- df %>% 
    group_by(Accepted_species) %>% 
    summarize(record_count = n()) %>% 
    filter(record_count > 3) %>% 
    nrow()
  
  # Print the results
  cat("Total number of species:", total_species, "\n")
  cat("Number of species with only one trait record:", species_with_one_record, "\n")
  cat("Number of species with more than three records:", species_with_more_than_three_records, "\n")
}
```



## Dispersal syndrome
```{r}
traits_dispersal <- plant_traits %>%
  filter(grepl('Dispersal syndrome|whole plant dispersal syndrome|dispersal_syndrome_1|dispersal_syndrome_2', TraitName))
traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
nrow(traits_dispersal)
length(unique(traits_dispersal$Accepted_species))
```

```{r}
# remove rows containing weight or mass
traits_dispersal <- traits_dispersal[!grepl("weight|mass|air-dry|oven|mean|SE",traits_dispersal$TraitValue, ignore.case = TRUE),]
nrow(traits_dispersal)
length(unique(traits_dispersal$Accepted_species))
```

```{r}
# remove rows where TraitValue is numeric
traits_dispersal <- traits_dispersal %>%
  filter(is.na(as.numeric(TraitValue)))
nrow(traits_dispersal)
length(unique(traits_dispersal$Accepted_species))
```


```{r}
traits_dispersal %>%
  distinct(TraitValue)
```

Reclassify to the following:
anemochorous - wind dispersal
anthropochorous - human dispersal
autochorous - non-aided dispersal
hydrochorous - water dispersal
myrmecochorous - ant dispersal
zoochorous - animal dispersal
unspecialized - multiple
```{r}
traits_dispersal <- traits_dispersal %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Eaten|Animal|Bird|Mammal|animals|endozoochor|dysochor|epizoochor|zoochor|Vertebrates|Reptiles+birds|Endo-zoochory|mammals|exozoochory|birds|Zoochory|vertebrate|Dys-zoochory|Ornithochory|Epi-zoochory|endozoochorous|endozoochory|ectozoochorous|zoochorous', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
    grepl('Ants', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
    grepl('wind|Anemochory|anemochor|meteorochor|chamaechor|Anemochory|anemochorous', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
    grepl('human|agochor|man|hemerochor|anthropochorous', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
    grepl('No|Unassisted|autochor|blastochor|ombrochor|Autochory|Barochory|ballochor|herpochor|autochorous', OrigTraitValue, ignore.case = TRUE) ~ "autochorous",
    grepl('water|nautochor|floating|hydrochor|hydrochorous', OrigTraitValue, ignore.case = TRUE) ~ "hydrochorous",
    grepl('boleochor|combination|non specialized|unspecialized', OrigTraitValue, ignore.case = TRUE) ~ "unspecialized"
  )) %>%
  filter(!is.na(TraitValue))
nrow(traits_dispersal)
length(unique(traits_dispersal$Accepted_species))
```

```{r}
traits_dispersal %>%
  distinct(TraitValue)
```

```{r}
# standardize trait names
traits_dispersal <- traits_dispersal %>%
  mutate(TraitName = "DispersalSyndrome")
```

```{r}
summarize_species_records(traits_dispersal)
```

```{r}
averagetraits_dispersal <- factor_data_merge(traits_dispersal, "DispersalSyndrome")
```

```{r}
nrow(averagetraits_dispersal)
length(unique(averagetraits_dispersal$Accepted_species))
```

```{r}
averagetraits_dispersal %>%
  count(TraitValue)
```

```{r}
summary(averagetraits_dispersal)
```


## Fruit dry mass
```{r}
traits_fruitMass <- plant_traits %>%
  filter(grepl('Fruit dry mass', TraitName))
traits_fruitMass$OrigTraitValue <- traits_fruitMass$TraitValue
nrow(traits_fruitMass)
length(unique(traits_fruitMass$Accepted_species))
```

```{r}
# remove non-numeric rows
traits_fruitMass <- traits_fruitMass %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_fruitMass)
length(unique(traits_fruitMass$Accepted_species))
```

```{r}
# check that units match
print(unique(traits_fruitMass$Unit))
```

```{r}
# standardize trait names with units
traits_fruitMass <- traits_fruitMass %>%
  mutate(TraitName = "FruitMass_mg")
```

```{r}
dotchart(traits_fruitMass$TraitValue)
```

```{r}
# detect outliers
fruitMass_outliers <- detect_outliers(traits_fruitMass)
print(fruitMass_outliers)
fruitMass_outliers_species <- species_with_outliers(fruitMass_outliers)
print(fruitMass_outliers_species)
```

```{r}
summarize_species_records(traits_fruitMass)
```


```{r}
averagetraits_fruitMass <- numeric_data_average(traits_fruitMass, "mg", "FruitMass_mg")
```

```{r}
nrow(averagetraits_fruitMass)
length(unique(averagetraits_fruitMass$Accepted_species))
```

```{r}
summary(averagetraits_fruitMass)
```


## Fruit length
```{r}
traits_fruitLength <- plant_traits %>%
  filter(grepl('Fruit length|maximum fruit length|minimum fruit length|fruit_length_min|fruit_length_max|fruit_length_mean', TraitName))
traits_fruitLength$OrigTraitValue <- traits_fruitLength$TraitValue

nrow(traits_fruitLength)
length(unique(traits_fruitLength$Accepted_species))
```

```{r}
# remove rows if TraitValue is non-numeric
traits_fruitLength <- traits_fruitLength %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_fruitLength)
length(unique(traits_fruitLength$Accepted_species))
```

```{r}
# check that units match
print(unique(traits_fruitLength$Unit))
```

```{r}
# if units = cm, multiply trait value by 10 and change units to mm
traits_fruitLength <- traits_fruitLength %>%
  mutate(TraitValue = if_else(Unit == "cm", TraitValue * 10, TraitValue),
         Unit = if_else(Unit == "cm", "mm", Unit))
```

```{r}
# standardize trait names with units
traits_fruitLength <- traits_fruitLength %>%
  mutate(TraitName = "FruitLength_mm")
```

```{r}
dotchart(traits_fruitLength$TraitValue)
```

```{r}
# detect outliers
fruitLength_outliers <- detect_outliers(traits_fruitLength)
print(fruitLength_outliers)
fruitLength_outliers_species <- species_with_outliers(fruitLength_outliers)
print(fruitLength_outliers_species)
```

```{r}
summarize_species_records(traits_fruitLength)
```

```{r}
averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "FruitLength_mm")
```

```{r}
nrow(averagetraits_fruitLength)
length(unique(averagetraits_fruitLength$Accepted_species))
```

```{r}
summary(averagetraits_fruitLength)
```


## Plant height
```{r}
traits_plantHeight <- plant_traits %>%
  filter(grepl('Plant height generative|Plant height vegetative|maximum whole plant height|minimum whole plant height|plant_height|plant_height_max|plant_height_min|whole plant height', TraitName))
traits_plantHeight$OrigTraitValue <- traits_plantHeight$TraitValue

nrow(traits_plantHeight)
length(unique(traits_plantHeight$Accepted_species))
```

```{r}
# remove non-numeric rows
traits_plantHeight <- traits_plantHeight %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_plantHeight)
length(unique(traits_plantHeight$Accepted_species))
```

```{r}
print(unique(traits_plantHeight$Unit))
```

```{r}
# standardize trait names with units
traits_plantHeight <- traits_plantHeight %>%
  mutate(TraitName = "PlantHeight_m")
```

```{r}
dotchart(traits_plantHeight$TraitValue)
```

```{r}
# detect outliers
plantHeight_outliers <- detect_outliers(traits_plantHeight)
print(plantHeight_outliers)
plantHeight_outliers_species <- species_with_outliers(plantHeight_outliers)
print(plantHeight_outliers_species)
```

```{r}
summarize_species_records(traits_plantHeight)
```

```{r}
averagetraits_plantHeight <- numeric_data_average(traits_plantHeight, "m", "PlantHeight_m")
```

```{r}
nrow(averagetraits_plantHeight)
length(unique(averagetraits_plantHeight$Accepted_species))
```


```{r}
summary(averagetraits_plantHeight)
```


## Plant lifespan/longevity
```{r}
traits_plantLifespan <- traits %>%
  filter(grepl('Plant lifespan \\(longevity\\)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
traits_plantLifespan$OrigTraitValue <- traits_plantLifespan$TraitValue

nrow(traits_plantLifespan)
length(unique(traits_plantLifespan$Accepted_species))
```

```{r}
# remove rows if TraitValue is non-numeric
traits_plantLifespan <- traits_plantLifespan %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_plantLifespan)
length(unique(traits_plantLifespan$Accepted_species))
```

```{r}
# check that units match
print(unique(traits_plantLifespan$Unit))
traits_plantLifespan$Unit <- "years"
print(unique(traits_plantLifespan$Unit))
```

```{r}
# standardize trait names with units
traits_plantLifespan <- traits_plantLifespan %>%
  mutate(TraitName = "PlantLifespan_years")

```

```{r}
dotchart(traits_plantLifespan$TraitValue)
```

```{r}
# detect outliers
plantLifespan_outliers <- detect_outliers(traits_plantLifespan)
print(plantLifespan_outliers)
plantLifespan_outliers_species <- species_with_outliers(plantLifespan_outliers)
print(plantLifespan_outliers_species)
```

```{r}
summarize_species_records(traits_plantLifespan)
```

```{r}
averagetraits_plantLifespan <- numeric_data_average(traits_plantLifespan, "years", "PlantLifespan_years")
```

```{r}
nrow(averagetraits_plantLifespan)
length(unique(averagetraits_plantLifespan$Accepted_species))
```

```{r}
summary(averagetraits_plantLifespan)
```

## Seed mass
```{r}
traits_seedMass <- traits %>%
  filter(grepl('Seed dry mass|seed mass|seed_mass_max|seed_mass_mean|seed_mass_min', TraitName))
traits_seedMass$OrigTraitValue <- traits_seedMass$TraitValue

nrow(traits_seedMass)
length(unique(traits_seedMass$Accepted_species))
```

```{r}
# remove non-numeric rows
traits_seedMass <- traits_seedMass %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_seedMass)
length(unique(traits_seedMass$Accepted_species))
```

```{r}
print(unique(traits_seedMass$Unit))
```

```{r}
# convert rows with unit = g to mg; multiply by 1000
traits_seedMass <- traits_seedMass %>%
  mutate(TraitValue = if_else(Unit == "g", TraitValue * 1000, TraitValue),
         Unit = if_else(Unit == "g", "mg", Unit))
```

```{r}
# standardize trait names with units
traits_seedMass <- traits_seedMass %>%
  mutate(TraitName = "SeedMass_g")
```

```{r}
dotchart(traits_seedMass$TraitValue)
```

```{r}
# detect outliers
seedMass_outliers <- detect_outliers(traits_seedMass)
print(seedMass_outliers)
seedMass_outliers_species <- species_with_outliers(seedMass_outliers)
print(seedMass_outliers_species)
```

```{r}
summarize_species_records(traits_seedMass)
```

```{r}
averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "SeedMass_g")
```

```{r}
nrow(averagetraits_seedMass)
length(unique(averagetraits_seedMass$Accepted_species))
```

```{r}
summary(averagetraits_seedMass)
```


## Seed length
```{r}
traits_seedLength <- traits %>%
  filter(grepl('Seed length|seed_length_max|seed_length_mean|seed_length_min', TraitName))
traits_seedLength$OrigTraitValue <- traits_seedLength$TraitValue

nrow(traits_seedLength)
length(unique(traits_seedLength$Accepted_species))
```

```{r}
# remove non-numeric rows
traits_seedLength <- traits_seedLength %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_seedLength)
length(unique(traits_seedLength$Accepted_species))
```

```{r}
print(unique(traits_seedLength$Unit))
```

```{r}
# standardize trait names with units
traits_seedLength <- traits_seedLength %>%
  mutate(TraitName = "SeedLength_mm")
```

```{r}
dotchart(traits_seedLength$TraitValue)
```

```{r}
# detect outliers
seedLength_outliers <- detect_outliers(traits_seedLength)
print(seedLength_outliers)
seedLength_outliers_species <- species_with_outliers(seedLength_outliers)
print(seedLength_outliers_species)
```

```{r}
summarize_species_records(traits_seedLength)
```

```{r}
averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "SeedLength_mm")
```

```{r}
nrow(averagetraits_seedLength)
length(unique(averagetraits_seedLength$Accepted_species))
```

```{r}
summary(averagetraits_seedLength)
```


## Seed width
```{r}
traits_seedWidth <- traits %>%
  filter(grepl('Seed width|seed_width_max|seed_width_mean|seed_width_min'
, TraitName))
traits_seedWidth$OrigTraitValue <- traits_seedWidth$TraitValue

nrow(traits_seedWidth)
length(unique(traits_seedWidth$Accepted_species))
```

```{r}
# remove non-numeric rows
traits_seedWidth <- traits_seedWidth %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))

nrow(traits_seedWidth)
length(unique(traits_seedWidth$Accepted_species))
```


```{r}
print(unique(traits_seedWidth$Unit))
```

```{r}
# standardize trait names with units
traits_seedWidth <- traits_seedWidth %>%
  mutate(TraitName = "SeedWidth_mm")
```

```{r}
dotchart(traits_seedWidth$TraitValue)
```

```{r}
# detect outliers
seedWidth_outliers <- detect_outliers(traits_seedWidth)
print(seedWidth_outliers)
seedWidth_outliers_species <- species_with_outliers(seedWidth_outliers)
print(seedWidth_outliers_species)
```

```{r}
summarize_species_records(traits_seedWidth)
```

```{r}
averagetraits_seedWidth <- numeric_data_average(traits_seedWidth, "mm", "SeedWidth_mm")
```

```{r}
nrow(averagetraits_seedWidth)
length(unique(averagetraits_seedWidth$Accepted_species))
```

```{r}
summary(averagetraits_seedWidth)
```


## Growth form
Just using data from GIFT 
```{r}
traits_growthForm <- plant_traits %>%
  filter(TraitName == 'growth_form')
traits_growthForm$OrigTraitValue <- traits_growthForm$TraitValue

nrow(traits_growthForm)
length(unique(traits_growthForm$Accepted_species))
```

```{r}
# remove numeric rows
traits_growthForm <- traits_growthForm %>%
  filter(is.na(as.numeric(TraitValue)))

nrow(traits_growthForm)
length(unique(traits_growthForm$Accepted_species))
```

```{r}
# look at unique trait values
traits_growthForm %>%
  distinct(TraitValue)
```

```{r}
# standardize trait names with units
traits_growthForm <- traits_growthForm %>%
  mutate(TraitName = "GrowthForm")
```

```{r}
summarize_species_records(traits_growthForm)
```

```{r}
averagetraits_growthForm <- factor_data_merge(traits_growthForm, "GrowthForm")
```

```{r}
nrow(averagetraits_growthForm)
length(unique(averagetraits_growthForm$Accepted_species))
```

```{r}
summary(averagetraits_growthForm)
```

# Fruit type
```{r}
traits_fruitType <- plant_traits %>%
  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue

nrow(traits_fruitType)
length(unique(traits_fruitType$Accepted_species))
```

```{r}
# remove numeric values
traits_fruitType <- traits_fruitType %>%
  filter(is.na(as.numeric(TraitValue)))

nrow(traits_fruitType)
length(unique(traits_fruitType$Accepted_species))
```

```{r}
traits_fruitType %>%
  count(TraitValue)
```

```{r}
# standardize trait values
traits_fruitType <- traits_fruitType %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
    grepl('capsule|Capsule|Capsule |Pixidium|Dehiscent capsule|Indehiscent capsule', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
    grepl('drupe|Drupe|Drupe |Drupaceous', OrigTraitValue, ignore.case = TRUE) ~ "drupe",
    grepl('aggregate berries', OrigTraitValue, ignore.case = TRUE) ~ "aggregate berries",
    grepl('aggregate drupelets', OrigTraitValue, ignore.case = TRUE) ~ "aggregate drupelets",
    grepl('fleshy', OrigTraitValue, ignore.case = TRUE) ~ "fleshy",
    grepl('pome', OrigTraitValue, ignore.case = TRUE) ~ "pome",
    grepl('Squash', OrigTraitValue, ignore.case = TRUE) ~ "squash",
    grepl('aggregate follicles', OrigTraitValue, ignore.case = TRUE) ~ "aggregate follicles",
    grepl('aggregate nutlets', OrigTraitValue, ignore.case = TRUE) ~ "aggregate nutlets",
    grepl('achene', OrigTraitValue, ignore.case = TRUE) ~ "achene",
    grepl('follicle|Follicle ', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
    grepl('legume|Legume ', OrigTraitValue, ignore.case = TRUE) ~ "legume",
    grepl('lomentum', OrigTraitValue, ignore.case = TRUE) ~ "lomentum",
    grepl('nut', OrigTraitValue, ignore.case = TRUE) ~ "nut",
    grepl('pod|Pod|Pod ', OrigTraitValue, ignore.case = TRUE) ~ "pod",
    grepl('Samara|Samaroid', OrigTraitValue, ignore.case = TRUE) ~ "samara",
    grepl('schizocarp', OrigTraitValue, ignore.case = TRUE) ~ "schizocarp",
    grepl('siliqua', OrigTraitValue, ignore.case = TRUE) ~ "siliqua",
    grepl('syncarpous', OrigTraitValue, ignore.case = TRUE) ~ "syncarpous",
    grepl('utricle', OrigTraitValue, ignore.case = TRUE) ~ "utricle",
    grepl('other|apocarpous|pericarp, aril, arillode|pseudosyncarpous|vegetative|dry|fleshy', OrigTraitValue, ignore.case = TRUE) ~ "unspecified",
  )) %>%
  filter(!is.na(TraitValue))
```

```{r}
traits_fruitType %>%
  count(TraitValue)
```

```{r}
# standardize trait name
traits_fruitType <- traits_fruitType %>%
  mutate(TraitName = "FruitType")
```

```{r}
summarize_species_records(traits_fruitType)
```

```{r}
averagetraits_fruitType <- factor_data_merge(traits_fruitType, "FruitType")
```

```{r}
nrow(averagetraits_fruitType)
length(unique(averagetraits_fruitType$Accepted_species))
```

```{r}
summary(averagetraits_fruitType)
```

# Combine all traits
```{r}
trait_dfs <- list(averagetraits_fruitLength, averagetraits_fruitMass, averagetraits_plantHeight, averagetraits_plantLifespan, averagetraits_seedLength, averagetraits_seedMass, averagetraits_seedWidth, averagetraits_dispersal, averagetraits_growthForm, averagetraits_fruitType)

# Function to combine trait dataframes only matching columns
combine_matching_columns <- function(df_list) {
  # Find the common columns across all data frames
  common_cols <- Reduce(intersect, lapply(df_list, colnames))
  
  # Ensure TraitValue column is of the same type across all data frames
  df_list <- lapply(df_list, function(df) {
    if ("TraitValue" %in% colnames(df)) {
      df <- df %>% mutate(TraitValue = as.character(TraitValue))
    }
    return(df)
  })
  
  # Filter each data frame to keep only the common columns
  df_list_filtered <- lapply(df_list, function(df) {
    df %>% select(all_of(common_cols))
  })
  
  # Combine the data frames using reduce and full_join
  combined_df <- reduce(df_list_filtered, full_join, by = common_cols)
  
  return(combined_df)
}
clean_traits <- combine_matching_columns(trait_dfs)
nrow(clean_traits)
```

```{r}
clean_traits %>%
  count(TraitName)
```

# Transform plant dataframe from long to wide
```{r}
wide_traits <- spread(clean_traits, key = TraitName, value = TraitValue)

#remove X column from dataframe
wide_plant_traits <- wide_traits[, !colnames(wide_traits) %in% c("X", "records_used", "variance"), drop = FALSE]
```

```{r}
# Assign data types
wide_plant_traits$`DispersalSyndrome` <- as.factor(wide_plant_traits$`DispersalSyndrome`)
wide_plant_traits$`FruitLength_mm` <- as.numeric(wide_plant_traits$`FruitLength_mm`)
wide_plant_traits$`FruitMass_mg` <- as.numeric(wide_plant_traits$`FruitMass_mg`)
wide_plant_traits$`FruitType` <- as.factor(wide_plant_traits$`FruitType`)
wide_plant_traits$`GrowthForm` <- as.factor(wide_plant_traits$`GrowthForm`)
wide_plant_traits$`PlantHeight_m` <- as.numeric(wide_plant_traits$`PlantHeight_m`)
wide_plant_traits$`PlantLifespan_years` <- as.numeric(wide_plant_traits$`PlantLifespan_years`)
wide_plant_traits$`SeedLength_mm` <- as.numeric(wide_plant_traits$`SeedLength_mm`)
wide_plant_traits$`SeedMass_g` <- as.numeric(wide_plant_traits$`SeedMass_g`)
wide_plant_traits$`SeedWidth_mm` <- as.numeric(wide_plant_traits$`SeedWidth_mm`)
```

```{r}
# rename accepted_species column to species
wide_plant_traits <- wide_plant_traits %>%
  rename(species = Accepted_species)
```


# Summary
```{r}
cat("Number of records:", nrow(clean_traits), "\n")
cat("Number of species:", length(unique(clean_traits$Accepted_species)), "\n")
```

```{r}
fb_plot_species_traits_completeness(wide_plant_traits)
fb_plot_number_species_by_trait(wide_plant_traits)
fb_table_trait_summary(wide_plant_traits)
```

```{r}
vis_dat(wide_plant_traits)
vis_miss(wide_plant_traits)
```


# Write data to csv
```{r}
write.csv(wide_plant_traits, file.path(output_path,"TropicalAndes_all_plant_traits_standardized.csv"))
write.csv(clean_traits, file.path(output_path,"TropicalAndes_all_plant_traits_standardized_with_recordCount.csv"))
```