---
title: "Tropical Andes Taxonomic Diversity of Plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Calculating and mapping taxonomic diversity as species richness using occurrence data for plants and frugivores in the Tropical Andes Moist Lowland and Montane forests. Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
date: "2023-08-01"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```


# Convert data to spatial data
```{r}
plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

# Transform to projected coordinate reference system (units from degrees to meters)
```{r}
Americas <- st_transform(Americas, 5389)
TApoly <- st_transform(TApoly, 5389)
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
plants.sf <- st_transform(plants.sf, 5389)
frugivores.sf <- st_transform(frugivores.sf, 5389)
```

# Group by species
```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoresPointsPlot
```


# Plot TD/Richness

## 5 km resolution
### Make Grid
```{r}
#check units
st_crs(TApoly, parameters = TRUE)$units_gdal
st_crs(TApoly)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

```{r}
TAGrid_5 <- TApoly %>%
  st_make_grid(cellsize = c(5000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_5) +
  labs(title = "Tropical Andes Grid [5 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_5 <- TAGrid_5 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_5 <- TAGrid_5 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_5km <-
ggplot(plant_richness_grid_5) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [5 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_5km
```

```{r}
frugivoregridRichTA_5km <-
ggplot(frugivore_richness_grid_5) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [5 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_5km
```


## Extract cell values
```{r}
plant_cellRichness_5km <- plantgridRichTA_5km[["data"]]
frugivore_cellRichness_5km <- frugivoregridRichTA_5km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_5km <- summary(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
compare_richness_summary_5km

# save r-squared value
compare_richness_r2_5km <- compare_richness_summary_5km$r.squared
```

```{r}
plot(plant_cellRichness_5km$num_species, frugivore_cellRichness_5km$num_species,
     main = "Plant Richness vs Frugivore Richness [5 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_5km$num_species ~ plant_cellRichness_5km$num_species))
```

```{r}
hist(plant_cellRichness_5km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_5km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

## 10 km
ADD CODE HERE!!!

## 25 km 
### Make Grid
```{r}
TAGrid_25 <- TApoly %>%
  st_make_grid(cellsize = c(25000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_25) +
  labs(title = "Tropical Andes Grid [25 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_25 <- TAGrid_25 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_25 <- TAGrid_25 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_25km <-
ggplot(plant_richness_grid_25) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [25 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_25km
```

```{r}
frugivoregridRichTA_25km <-
ggplot(frugivore_richness_grid_25) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [25 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_25km
```


## Extract cell values
```{r}
plant_cellRichness_25km <- plantgridRichTA_25km[["data"]]
frugivore_cellRichness_25km <- frugivoregridRichTA_25km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_25km <- summary(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
compare_richness_summary_25km

# save r-squared value
compare_richness_r2_25km <- compare_richness_summary_25km$r.squared
```

```{r}
plot(plant_cellRichness_25km$num_species, frugivore_cellRichness_25km$num_species,
     main = "Plant Richness vs Frugivore Richness [25 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_25km$num_species ~ plant_cellRichness_25km$num_species))
```

```{r}
hist(plant_cellRichness_25km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_25km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```


## 50 km 
### Make Grid
```{r}
TAGrid_50 <- TApoly %>%
  st_make_grid(cellsize = c(50000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_50) +
  labs(title = "Tropical Andes Grid [50 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_50 <- TAGrid_50 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_50 <- TAGrid_50 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_50km <-
ggplot(plant_richness_grid_50) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [50 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_50km
```

```{r}
frugivoregridRichTA_50km <-
ggplot(frugivore_richness_grid_50) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [50 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_50km
```

## Extract cell values
```{r}
plant_cellRichness_50km <- plantgridRichTA_50km[["data"]]
frugivore_cellRichness_50km <- frugivoregridRichTA_50km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_50km <- summary(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
compare_richness_summary_50km

# save r-squared value
compare_richness_r2_50km <- compare_richness_summary_50km$r.squared
```

```{r}
plot(plant_cellRichness_50km$num_species, frugivore_cellRichness_50km$num_species,
     main = "Plant Richness vs Frugivore Richness [50 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_50km$num_species ~ plant_cellRichness_50km$num_species))
```

```{r}
hist(plant_cellRichness_50km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_50km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```


## 75 km 
### Make Grid
```{r}
TAGrid_75 <- TApoly %>%
  st_make_grid(cellsize = c(75000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_75) +
  labs(title = "Tropical Andes Grid [75 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_75 <- TAGrid_75 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_75 <- TAGrid_75 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_75km <-
ggplot(plant_richness_grid_75) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [75 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_75km
```

```{r}
frugivoregridRichTA_75km <-
ggplot(frugivore_richness_grid_75) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [75 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_75km
```

## Extract cell values
```{r}
plant_cellRichness_75km <- plantgridRichTA_75km[["data"]]
frugivore_cellRichness_75km <- frugivoregridRichTA_75km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_75km <- summary(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
compare_richness_summary_75km

# save r-squared value
compare_richness_r2_75km <- compare_richness_summary_75km$r.squared
```

```{r}
plot(plant_cellRichness_75km$num_species, frugivore_cellRichness_75km$num_species,
     main = "Plant Richness vs Frugivore Richness [75 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_75km$num_species ~ plant_cellRichness_75km$num_species))
```

```{r}
hist(plant_cellRichness_75km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_75km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```

## 100 km 
### Make Grid
```{r}
TAGrid_100 <- TApoly %>%
  st_make_grid(cellsize = c(100000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())
```

### Plot Empty Grid
```{r}
gridPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest) +
  geom_sf(data = TAGrid_100) +
  labs(title = "Tropical Andes Grid [100 km]", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

gridPlot
```

### Cell Richness
```{r}
plant_richness_grid_100 <- TAGrid_100 %>%
  st_join(plants_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

```{r}
frugivore_richness_grid_100 <- TAGrid_100 %>%
  st_join(frugivores_sf_species) %>%
  mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
  group_by(cellid) %>%
  summarize(num_species = sum(overlap))
```

### Richness plots
```{r}
plantgridRichTA_100km <-
ggplot(plant_richness_grid_100) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Plant Richness [100 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantgridRichTA_100km
```

```{r}
frugivoregridRichTA_100km <-
ggplot(frugivore_richness_grid_100) +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_sf(aes(fill = num_species), color = NA) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  labs(title = "Tropical Andes Frugivore Richness [100 km]", x = "Longitude", y = "Latitude",
       fill = "Number of species") +
    coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoregridRichTA_100km
```

## Extract cell values
```{r}
plant_cellRichness_100km <- plantgridRichTA_100km[["data"]]
frugivore_cellRichness_100km <- frugivoregridRichTA_100km[["data"]]
```

### Plant vs frugivore richness plot
```{r}
compare_richness_summary_100km <- summary(lm(frugivore_cellRichness_100km$num_species ~ plant_cellRichness_100km$num_species))
compare_richness_summary_100km

# save r-squared value
compare_richness_r2_100km <- compare_richness_summary_100km$r.squared
```

```{r}
plot(plant_cellRichness_100km$num_species, frugivore_cellRichness_100km$num_species,
     main = "Plant Richness vs Frugivore Richness [100 km]",
     xlab = "Plant Richness by cell", ylab = "Frugivore Richness by cell")
abline(lm(frugivore_cellRichness_100km$num_species ~ plant_cellRichness_100km$num_species))
```

```{r}
hist(plant_cellRichness_100km$num_species,
     main = "Histogram of Plant Richness by cell",
     xlab = "Plant Richness by cell")
hist(frugivore_cellRichness_100km$num_species,
     main = "Histogram of Frugivore Richness by cell",
     xlab = "Frugivore Richness by cell")
```






# Compare r-squared by resolution
Plot
```{r}
# create dataframe of resolutions and r-squared values
resolution <- c("5", "25", "50", "75", "100")
r_squared <- c(compare_richness_r2_5km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
r_squared_compare <- data.frame(resolution, r-squared)
```

```{r}
ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
  geom_point()+
  labs(title = "R-squared values of plant-frugivore richness relationships by resolution", x = "Resolution (km)", y = "R-squared",
       fill = "Number of species") 
```


# Multi panel plots
Richness maps
```{r}
grid.arrange(plot_1, plot_2, nrow = 1)
```

# Write data to csv
```{r}
write.csv(plant_cellRichness0.75, file.path(output_path,"TropicalAndes_GBIF_plant_richness_sub.csv"))
write.csv(frugivore_cellRichness0.75, file.path(output_path,"TropicalAndes_frugivore_richness_sub.csv"))
```
