---
title: "Tropical Andes plant trait data GIFT"
author: "Hazel J. Anderson, Jenna B. Baljunas"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the GIFT database for plant species list."
data input: "none"
data output: "TropicalAndes_GIFT_plant_traits.csv"
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
```

# Retrive GIFT data
Data downloaded 2023-10-25
```{r}
all_traits <- GIFT_traits_meta()
```
Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3), Plant_height_min (1.6.1), Plant_height_max (1.6.2), Fruit_type_1 (3.16.1), fruit_dryness_1 (3.18.1), Fruit_length_min (3.13.1), Fruit_length_max (3.13.2), Fruit_length_mean (3.13.3), Seed_mass_min (3.2.1), Seed_mass_max (3.2.2), Seed_mass_mean (3.2.3), Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3), Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3), Growth_form (1.2.1)

```{r}
# list all the traits of interest here:
ID <- c('1.6.3','1.6.1','1.6.2','2.2.1','3.3.1','3.3.2','3.16.1','3.18.1','3.13.3','3.13.1','3.13.2','3.2.3','3.2.1','3.2.2','3.10.3','3.10.1','3.10.2','3.11.3','3.11.1','3.11.2','1.2.1')

traits <- lapply(ID,function(x){
  # extract row of each ID from entire list of traits 
  desired_trait <- all_traits[all_traits$Lvl3 == x,]
  
  # import GIFT data for each trait
  GIFT_data <- GIFT_traits(trait_IDs=x)
  
  # only keep species and trait value columns (consistently columns 2 and 4)
  GIFT_data <- GIFT_data[,c(2,4)]
  
  # rename these columns for combining later
  colnames(GIFT_data)<- c('species','trait_value')
  
  # characterize values
  GIFT_data$trait_value <- as.character(GIFT_data$trait_value)
  
  # add trait name column
  GIFT_data$trait_name <- desired_trait$Trait2
  
  # create separate df
  df <- data.frame(species=GIFT_data$species, trait_value=GIFT_data$trait_value, trait_name=GIFT_data$trait_name)
})

# data for all desired traits (called 2 for now to compare to original data):  
GIFT_data2 <- bind_rows(traits)
```

```{r}
glimpse(GIFT_data2)
```

```{r}
paste("The number of records is", nrow(GIFT_data2))
paste("The number of species is", length(unique(GIFT_data2$species)))
```

```{r}
GIFT_data2 %>% count(trait_name)
```

# Write data to csv
```{r}
write.csv(GIFT_data2, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits2.csv"))
```